Leaf traits drive differences in biomass partitioning among major plant functional types

### 2015-07-21 manuscript.Rnw
More writing!; Formatting is now complete, I think.

Perhaps the biggest challenge for GVMs is to capture the combined responses of the more than 250,000 plant species comprising the world's vegetation. While most plants have the same basic resource requirements and physiological function, large differences exist among species in the amount of energy invested in different tissues (leaves, stems, roots). The approach taken by most GVMs for dealing with this functional diversity is to consider only a few archetypal plant functional types (PFTs) \citep{harrisonetal2010}, assumed to differ in key physiological attributes. While GVMs assume or predict differences in biomass partitioning between PFTs (Methods \ref{sec:gvmsummary}), these differences are constrained by limited empirical data. Moreover, there is little consensus on how biomass partitioning and allocation (see Methods \ref{sec:partterm} for terminology) should be modelled in GVMs \citep{franklin2012,dekauwe2014,friend2014}. These shortcomings largely reflect the lack of suitable datasets of global scope with which models can be tested, constrained and compared \citep{wolf2011}.

@@

Little is known about global-scale patterns in LMF and LAR in relation to climate. One may expect that biomass partitioning is correlated with precipitation or mean annual temperature \citep{reich2014} because short-term experiments (cite) and climate gradients \citep{callaway1994,poyatos2007} have shown responsiveness of biomass partitioning and related traits. It is thus the question whether variation within PFT can be explained by climate or not. Indeed, many GVMs assume that climate affects biomass partitioning within a given PFT, often through soil water stress or other abiotic stress factors (see Methods \ref{sec:gvmsummary}).

@@

We used the recently compiled Biomass And Allometry Database (BAAD) \citep{Falster2015}, including a total of \Sexpr{nrow(baad_all$data)} observations, but here we use only data that include leaf mass ($M_F$), leaf area ($A_F$), above-ground woody biomass ($M_S$), plant height ($H$), and stem area measured at ground level ($A_S$), or at breast height (ca. 1.3m) ($A_{Sbh}$).  For each analysis, we used different subsets of the data. Sample sizes by PFT are summarized in Tables~\ref{table:count1}-\ref{table:count4}. We excluded glasshouse and common garden studies, including only field-grown woody plants (including natural vegetation, unmanaged and managed plantations). We considered three PFTs : evergreen angiosperms, evergreen gymnosperms, and deciduous angiosperms. We excluded  deciduous gymnosperms because few data were available. All locations were further separated into boreal (including sub-boreal), temperate, and tropical biomes. To assess the coverage of the global climate space by the dataset, we extracted mean annual temperature and precipitation from Worldclim \citep{worldclim}, excluding areas without woody vegetation (taken from the global land cover database GLC-SHARE \citep{lathamglobal}).

@@

We only used LMA directly estimated for the harvested plants (typically for a subsample of leaves, see \citet{Falster2015} for methods details for each contributed study). For conifers, leaf area was converted to half-total surface area using the average of a set of published conversion factors \citep{barclay2000}. We used a different conversion for pines (\emph{Pinus} spp.) than non-pines, because the conversion factors were very different for these two groups \citep{barclay2000}. This conversion was necessary because half-total surface area is most appropriate for comparison to flat leaves \citep{lang1991, chen1992}. Stem cross-sectional area was measured either at breast height (ca. 1.3m) and/or at the base of the plant. For the subset of the data where both were measured, we estimated basal stem area ($A_S$) from breast height stem area ($A_{Sbh}$) using a non-linear regression model, as follows.

@@

$$
D_S = D_{Sbh} H^a / (H - H_{bh})^a
$$

@@

$$
a = c_0H^{c_1}
$$

@@

We used generalized additive models (GAM) to visualize the relationships between biomass and plant size variables, and to estimate variables and their confidence intervals such as LMF at a common plant height. In all fitted GAMs, we used species-dataset combination as a random effect. We used the \emph{mgcv} package in R \citep{wood2006}. All variables (except MAP and MAT) were log-transformed prior to analysis.

@@

Because of allometric relationships between the studied variables \citep{niklas1994}, there was a strong effect of plant height on LMF (Fig. 2a and S2). We corrected for plant height by estimating LMF at a common plant height, using a generalized additive model, a semi-parametric method that makes no assumption on the underlying functional relationships \citep{wood2006}. We then found that LMF at a common height was proportional to the leaf construction cost, as measured by leaf mass per area (LMA) (Fig. 2b). As a consequence, LAR was invariant between PFTs, because LAR = LMF / LMA (Fig. 2c and S3).

Another key metric governing vegetation behavior is the amount of leaf maintained per unit cross-sectional area of stem ($A_S$). This ratio is relevant because plant leaf mass is frequently estimated from $A_S$ with allometric equations \citep{shinozaki1964b, chave2005}.  Mirroring the results for LMF, we found that the amount of leaf mass per unit stem area ($M_F/A_S$) differs among the three PFTs, that these differences are also proportional to LMA (Fig. 2d), and that PFTs are similarly invariant in the amount of leaf area per unit stem area ($A_F/A_S$) (Fig. 2e).

The above results highlight how LMA drives differences in biomass partitioning among dominant woody PFTs. To assess whether LMA could replace PFT when predicting biomass partitioning, we fit a linear mixed-effects model to LMF and $M_F/A_S$ using PFT and plant height (and the quadratic term) as predictors (and all interactions). We then replaced PFT with LMA, and found that the model with LMA could explain almost as much variation in LMF as the model with PFT ($R^2$ = 0.74 with PFT vs. 0.62 with LMA) and $M_F/A_S$ ($R^2$ = 0.29 vs. 0.28). This is surprising because LMA only captures aspects of leaf morphology, in indirectly leaf lifespan and photosynthetic capacity via the LES\citep{wrightetal2004}. These results suggest promising avenues for parameterizing and simplifying biomass partitioning routines in GVMs.

@@

The correlation between $M_F/A_S$ and LMA was found to hold also within PFTs (Fig. 4a), explaining 30\% of the variation in $M_F/A_S$ across species. The regressions across species by PFT were broad and overlapping, with a more general relationship extending across the entire LMA axis. A larger fraction of the variation in $M_F/A_S$ was explained by species-level differences in $A_F/A_S$ (Fig. 4b). As previously noted, $A_F/A_S$ does not differ systematically among PFTs (Fig. 2e), but does vary close to two orders of magnitude across species. In Fig. 4b, The separation among PFTs in $M_F/A_S$ arises due to differences in LMA. We thus show that in a PFT-based classification, LMA is a good first estimate of biomass partitioning, however, more variation can be explained by another trait ($A_F/A_S$), which varies appreciably between species.

We found that decomposing as in Eq. \ref{eq:M1} was very useful to study climate effects because the second term absorbs nearly all of the size related variation in LMF (Tables \ref{table/varpartB} and \ref{table/varpartMAPMAT}) and is otherwise fairly conserved across PFTs (Fig. S4). As shown in Fig. 2b and Fig. 2d, the term $M_F/A_S$ exhibits comparable differences between PFTs to LMF, but unlike LMF, $M_F/A_S$ is nearly independent of plant height (Tables \ref{table/varpartB} and \ref{table/varpartMAPMAT}). $M_F/A_S$ is thus a useful proxy for LMF that can be compared across species and sites.

Regardless of how we analyzed the data, we found only weak climatic effects on biomass partitioning within PFTs (Fig. 3). Biome consistently explained very little variation when added to a statistical model in addition to PFT and plant height ($R^2$ increased by only 0.01 - 0.06, see Table \ref{table/varpartB}). Likewise, we found weak and inconsistent effects of climate variables (MAP and MAT) (Fig. 3b-c). These variables explained little variation when added to a statistical model in addition to PFT and plant height ($R^2$ increased by 0.01 and 0.11, respectively, see Table \ref{table/varpartMAPMAT}).

@@

We found that PFTs differed in LMF, and that LMF was proportional to LMA. This was true especially across the three PFTs studied, but also to a lesser extent across species within PFT. The consequence is that the amount of leaf area supported per unit biomass (leaf area ratio, LAR) did not differ between PFTs, and was much less variable within PFTs than LMF. This result is particularly convincing because we studied the entire size range of woody plants in natural forests (0.1-100m) (Fig. 2a). Other work has already demonstrated a large difference in LMF between angiosperms and gymnosperms, but have not linked this directly to differences in LMA, nor was able to tell if it is because of angiosperm-gymnosperm divide or because of deciduous vs. evergreen habit. We show that LMA -- a central trait of the leaf economics spectrum (LES)\citep{wrightetal2004} -- explains these differences in a consistent manner.

Poorter et al 2012 did not find a clear relationship between LMF and LMA across 8 PFTs, perhaps not enough data? Also, we only include field-grown plants, they have everything. It also seems they did not account for plant size in that analysis.

paragraph - Why would LAR be less variable than LMF? Because light interception. Lulu effect paper. Need to discuss linkage between LMA, leaf lifespan, leaf area vs. leaf mass at given MT (and given height!).
One option is that LAR is targeted to be more constant, because it drives light interception. Very high leaf areas lead to very shaded foliage (etc.). Other option is that this arises through the correlation with leaf lifepan. This is how Enquist \& Niklas explained it, but it is only true when all else is equal. In support, \citet{reich1992} showed positive correlation between leaf life span and total stand leaf biomass across diverse forest stands. Our results are certainly consistent with longer leaf lifespan leading to higher leaf biomass, because of the strong nearly 1/x correlation between LL and LMA ()

paragraph - variation in LMF was the same as MF/AS, suggesting that MT/AS is a fairly conserved quantity \ref{eq:M2}. Consistent with \citep{chave2005} who showed that aboveground biomass can be well estimated with c*Ht*AS*wood density (works because geometry). This works because, for large plants, most of aboveground biomass is wood.

Climate (temperature, rainfall) did not appreciably affect biomass partitioning between leaves and stems. This finding is in contrast with a recent study where LMF was found to increase with MAT across the globe \citep{reich2014}. One explanation for these differences is that our measurements were taken on individual plants, whereas in the other study they were on whole stands. This suggests climatic effects on stand-level biomass partitioning occur primarily by altering the size-distribution and stand density (number of individuals per unit area), rather than partitioning within individual plants. With the caveats that there is still variation between species and studies unaccounted for (Fig. S5), and that coarse climate variables such as MAT and MAP may mask small-scale climate effects on biomass partitioning, our results strongly suggest the effects of PFT are stronger than any direct climate effect on biomass partitioning within individual plants (Tables \ref{table/varpartB} and \ref{table/varpartMAPMAT}).

Overall, our results establish general patterns about plant construction and thus lay an empirical base against which models can be benchmarked. We found that PFTs differed in LMF, and that LMF was proportional to LMA. It thus seems possible to integrate a key leaf trait with whole-plant modelling of biomass partitioning. A recent study compared allocation routines in a number of leading ecosystem models \citep{dekauwe2014}, and recommended constraining allocation by observed biomass fractions instead of using constant allocation fractions. Based on our findings, a first approximation within GVMs would be to assume leaf area to stem cross-sectional area a parameter that does not vary between PFTs (notwithstanding effects of plant height on this variable), and vary biomass allocation accordingly. LMA, already a parameter in most GVMs, then gives the ratio of leaf biomass to stem area. Some models already incorporate a similar algorithm (see Methods~\ref{sec:gvmsummary}), while other algorithms may be tuned to yield similar patterns between PFTs as we have presented. In any case, the growing availability of large datasets on stand biomass and individual plant construction  \citep{Falster2015} suggest the time is ripe for rigorous benchmarking \citep{abramowitz_towards_2012, dekauwe2014} of GVMs against empirical data.

@@

>>>>

Perhaps the biggest challenge for GVMs is to capture the combined responses of the more than 250,000 plant species comprising the world's vegetation. While most plants have the same basic resource requirements and physiological function, large differences exist among species in the amount of energy invested in different tissues (leaves, stems, roots). The approach taken by most GVMs for dealing with this functional diversity is to consider only a few archetypal plant functional types (PFTs) \citep{harrisonetal2010}, assumed to differ in key physiological attributes. While GVMs assume or predict differences in biomass partitioning between PFTs (Notes S1), these differences are constrained by limited empirical data. Moreover, there is little consensus on how biomass partitioning and allocation (see Methods \ref{sec:partterm} for terminology) should be modelled in GVMs \citep{franklin2012,dekauwe2014,friend2014}. These shortcomings largely reflect the lack of suitable datasets of global scope with which models can be tested, constrained and compared \citep{wolf2011}.

@@

Little is known about global-scale patterns in LMF and LAR in relation to climate. It can be expected that biomass partitioning is correlated with precipitation or mean annual temperature \citep{reich2014} because short-term experiments (cite) and climate gradients \citep{callaway1994,poyatos2007} have shown responsiveness of biomass partitioning to environmental drivers. A recent study \citep{reich2014} also demonstrated that stand-scale biomass partitioning was related to mean annual temperature across diverse forest stands. Indeed, many GVMs assume that climate affects biomass partitioning within a given PFT, often through soil water stress or other abiotic stress factors (see Notes S1). Again, however, these models do not agree on the degree of plasticity in biomass partitioning, or which variables should it respond to.

@@

We used the recently compiled Biomass And Allometry Database (BAAD) \citep{Falster2015}, including a total of \Sexpr{nrow(baad_all$data)} observations, but here we use only data that include leaf mass ($M_F$), leaf area ($A_F$), above-ground woody biomass ($M_S$), plant height ($H$), and stem area measured at ground level ($A_S$), or at breast height (typically 1.3m) ($A_{Sbh}$).  For each analysis, we used different subsets of the data. Sample sizes by PFT are summarized in Tables S2-S5. We excluded glasshouse and common garden studies, including only field-grown woody plants (including natural vegetation, unmanaged and managed plantations). We considered three PFTs : evergreen angiosperms, evergreen gymnosperms, and deciduous angiosperms. We excluded  deciduous gymnosperms because few data were available. All locations were further separated into boreal (including sub-boreal), temperate, and tropical biomes. To assess the coverage of the global climate space by the dataset, we extracted mean annual temperature and precipitation from Worldclim \citep{worldclim}, excluding areas without woody vegetation (taken from the global land cover database GLC-SHARE \citep{lathamglobal}).

@@

We only used LMA directly estimated for the harvested plants (typically for a subsample of leaves, see \citet{Falster2015} for methods details for each contributed study). For conifers, leaf area was converted to half-total surface area using the average of a set of published conversion factors \citep{barclay2000}. We used a different conversion for pines (\emph{Pinus} spp.) than non-pines, because the conversion factors were very different for these two groups \citep{barclay2000}. This conversion was necessary because half-total surface area is most appropriate for comparison to flat leaves \citep{lang1991, chen1992}. Stem cross-sectional area was measured either at breast height and/or at the base of the plant. For the subset of the data where both were measured, we estimated basal stem area ($A_S$) from breast height stem area ($A_{Sbh}$) using a non-linear regression model, as follows.

@@

$$D_S = D_{Sbh} H^a / (H - H_{bh})^a$$

@@

$$a = c_0H^{c_1}$$

@@

We used generalized additive models (GAM) to visualize the relationships between biomass and plant size variables, and to estimate variables and their confidence intervals such as LMF at a common plant height. In all fitted GAMs, we used a cubic regression spline. For the smoothed term in the model (plant height), we used up to 3 or 4 degrees of freedom, which resulted in biologically realistic smoothed relationships. Within the GAM, we used a penalized regression smoother [@Wood2006] to allow the final degree of smoothness to be estimated from the data. In all fitted GAMs, we used species-dataset combination as a random effect.We used the \emph{mgcv} package in R \citep{wood2006}. All variables (except MAP and MAT) were log-transformed prior to analysis.

@@

Because of allometric relationships between the studied variables \citep{niklas1994}, there was a strong effect of plant height on LMF (Fig. 2a and S2). We corrected for plant height by estimating LMF at a common plant height, using a generalized additive model, a semi-parametric method that makes no assumption on the underlying functional relationships. We then found that LMF at a common height was proportional to the leaf construction cost, as measured by leaf mass per area (LMA) (Fig. 2b). As a consequence, LAR was invariant between PFTs (Eq. 1) (Fig. 2c and S3).

Another key metric governing vegetation behavior is the amount of leaf maintained per unit cross-sectional area of stem ($A_S$).  Mirroring the results for LMF, we found that the amount of leaf mass per unit stem area ($M_F/A_S$) differs among the three PFTs, that these differences are also proportional to LMA (Fig. 2d), and that PFTs are similarly invariant in the amount of leaf area per unit stem area ($A_F/A_S$) (Fig. 2e).

The above results highlight how LMA drives differences in biomass partitioning among dominant woody PFTs. To assess whether LMA could replace PFT when predicting biomass partitioning, we fit a linear mixed-effects model to LMF and $M_F/A_S$ using PFT and plant height (and the quadratic term) as predictors (and all interactions). We then replaced PFT with LMA, and found that the model with LMA could explain almost as much variation in LMF as the model with PFT ($R^2$ = 0.74 with PFT vs. 0.62 with LMA) and $M_F/A_S$ ($R^2$ = 0.29 vs. 0.28).

@@

The ratio $M_F/A_S$ is relevant because plant leaf mass is frequently estimated from $A_S$ with allometric equations \citep{shinozaki1964b, chave2005}. The correlation between $M_F/A_S$ and LMA was found to hold also within PFTs (Fig. 3a), explaining 30\% of the variation in $M_F/A_S$ across species. The regressions across species by PFT were broad and overlapping, with a more general relationship extending across the entire LMA axis. A larger fraction of the variation in $M_F/A_S$ was explained by species-level differences in $A_F/A_S$ (Fig. 3b). As previously noted, $A_F/A_S$ does not differ systematically among PFTs (Fig. 2e), but does vary close to two orders of magnitude across species. In Fig. 3b, The separation among PFTs in $M_F/A_S$ arises due to differences in LMA.

We found that decomposing as in Eq. \ref{eq:M1} was very useful to study climate effects because the second term absorbs nearly all of the size related variation in LMF (Tables \ref{table/varpartB} and S1) and is otherwise fairly conserved across PFTs (Fig. S4). As shown in Fig. 2b and Fig. 2d, the term $M_F/A_S$ exhibits comparable differences between PFTs to LMF, but unlike LMF, $M_F/A_S$ is nearly independent of plant height (Tables \ref{table/varpartB} and S1. $M_F/A_S$ is thus a useful proxy for LMF that can be compared across species and sites.

Regardless of how we analyzed the data, we found only weak climatic effects on biomass partitioning within PFTs (Fig. 4). Biome consistently explained very little variation when added to a statistical model in addition to PFT and plant height ($R^2$ increased by only 0.01 - 0.06, see Table \ref{table/varpartB}). Likewise, we found weak and inconsistent effects of climate variables (MAP and MAT) (Fig. 4b-c). These variables explained little variation when added to a statistical model in addition to PFT and plant height ($R^2$ increased by 0.01 and 0.11, respectively, see Table S1).

@@

We found that PFTs differed in LMF, and that LMF was proportional to LMA. This was true especially across the three PFTs studied, but also to a lesser extent across species within PFT. The consequence is that the amount of leaf area supported per unit biomass (leaf area ratio, LAR) did not differ between PFTs, and was much less variable within PFTs than LMF. This result is particularly convincing because we studied the entire size range of woody plants in natural forests (0.1-100m) (Fig. 2a). Previous studies have demonstrated a large difference in LMF between angiosperms and gymnosperms \citep{poorter2012}, but have not further separated functional types into deciduous vs. evergreen angiosperms, or provided an explanation in terms of leaf traits. We show that LMA -- a central trait of the leaf economics spectrum (LES)\citep{wrightetal2004} -- explains these differences in a consistent manner.

A previous large compilation of plant biomass partitioning \citep{poorter2012} (not restricted to field-grown plants) did not find a correlation between LMF and LMA across 8 PFTs (including non-woody PFTs). In their comparison, they did not correct for plant size, which has a very large effect on LMF (Fig. 2). It is certainly necessary to correct for plant size when comparing biomass partitioning parameters across PFTs.

The finding that LMF is proportional to LMA across PFTs implies that LAR does not differ. Why would this be the case? One explanation is that it is beneficial for plants to target a more constant LAR rather than LMF, because leaf area drives light interception. Because canopy size closely correlates with aboveground biomass \citep{duursma2010}, and the ratio of leaf area to canopy size determines self-shading and thus production efficiency of the foliage \citep{duursma20}, a low LAR is wasteful of woody biomass, and a high LAR wasteful of the foliage because self-shading becomes too high. Another, not mutually exclusive, explanation lies in the strong positive correlation between LMA and leaf lifespan (LL) \citep{wright2004}. Both \cite{poorter2012} and \citet{enquistniklas2002} explained higher LMF in gymnosperms by higher LL; plants simply maintain more foliage. However it is not not self-evident that else else is indeed equal, that is, annual production of foliage could differ between PFTs. Nonetheless, further supporting this explanation, \citet{reich1992} showed a positive correlation between LL and total stand leaf biomass across diverse forest stands.

We found that, as for LMF, $M_F/A_S$ was also proportional to LMA across PFTs (Fig. 2). Because LMF may be decomposed into the terms $M_F/A_S$ and $M_T/A_S$ (\ref{eq:M2}), this result suggests that $M_T/A_S$ does not vary between PFTs (see also Fig. S4). If this result holds, it means that aboveground biomass may be estimated across PFTs with a single equation. For tropical forests, \citet{chave2005} showed that aboveground biomass can be well estimated as,

$$M_T = F \rho A_S H_T$$

where F is a form factor, and $\rho$ wood density (kg m$^3$). In this equation leaf mass is subsumed in F, or assumed to be negligible (which hold for somewhat tall trees, see Fig. 2). Our result seems to indicate something like this works for all PFTs (see Fig. S4). However, there are differences in wood density between angiosperms and gymnosperms (Zanne + TRY categories), does this mean F is different between PFTs? It is timely to datamine BAAD to develop better general equations to estimate aboveground biomass.

Climate (temperature, rainfall) did not appreciably affect biomass partitioning between leaves and stems. This finding is in contrast with a recent study where LMF was found to increase with MAT across the globe \citep{reich2014}. One explanation for these differences is that our measurements were taken on individual plants, whereas in the other study they were on whole stands. This suggests climatic effects on stand-level biomass partitioning occur primarily by altering the size-distribution and stand density (number of individuals per unit area), rather than partitioning within individual plants. With the caveats that there is still variation between species and studies unaccounted for (Fig. S5), and that coarse climate variables such as MAT and MAP may mask small-scale climate effects on biomass partitioning, our results strongly suggest the effects of PFT are stronger than any direct climate effect on biomass partitioning within individual plants (Tables \ref{table/varpartB} and S1).

We found that PFTs differed in LMF, and that LMF was proportional to LMA. It thus seems possible to integrate a key leaf trait with whole-plant modelling of biomass partitioning. Indeed, nearly as much variation was explained across species when using LMA instead of PFT in a statistical model. This is surprising because LMA only captures aspects of leaf morphology, in indirectly leaf lifespan and photosynthetic capacity via the leaf economics spectrum \citep{wrightetal2004}. We thus show that in a PFT-based classification, LMA is a good first estimate of biomass partitioning, however, more variation can be explained by another trait ($A_F/A_S$), which varies appreciably between species. These results suggest promising avenues for parameterizing and simplifying biomass partitioning routines in GVMs.

Overall, our results establish general patterns about plant construction and thus lay an empirical base against which models can be benchmarked. A recent study compared allocation routines in a number of leading ecosystem models \citep{dekauwe2014}, and recommended constraining allocation by observed biomass fractions instead of using constant allocation fractions. Based on our findings, a first approximation within GVMs would be to assume leaf area to stem cross-sectional area a parameter that does not vary between PFTs (notwithstanding effects of plant height on this variable), and vary biomass allocation accordingly. LMA, already a parameter in most GVMs, then gives the ratio of leaf biomass to stem area. Some models already incorporate a similar algorithm (see Notes S1), while other algorithms may be tuned to yield similar patterns between PFTs as we have presented. In any case, the growing availability of large datasets on stand biomass and individual plant construction  \citep{Falster2015} suggest the time is ripe for rigorous benchmarking \citep{abramowitz_towards_2012, dekauwe2014} of GVMs against empirical data.

@@

\section*{Supporting Information}
Additional supporting information may be found in the online version of this article.\\

\textbf{Table S1} Explained variance by plant functional type (PFT), plant height (H), mean annual temperature (MAT) and mean annual precipitation (MAP), for five whole-plant variables.\\
\textbf{Table S2} Number of observations used in the analysis of $M_F/A_S$ and $A_F/A_S$.\\
\textbf{Table S3} Number of observations used in the analysis of $A_F/A_S$.\\
\textbf{Table S4} Number of observations used in the analysis of leaf mass fraction ($M_F/M_T$) and leaf area ratio ($A_F/M_T$).\\
\textbf{Table S5} Number of observations used in the analysis of leaf area ratio (LAR).\\
\textbf{Fig. S1} Global coverage of the climate space by the dataset, labelled by vegetation type.\\
\textbf{Fig. S2} Raw data for leaf biomass ($M_F$) and total above-ground woody biomass ($M_S$) for each of the PFTs, as a function of total plant height ($H$).\\
\textbf{Fig. S3} Leaf area ratio ($A_F/M_T$) by PFT.\\
\textbf{Fig. S4} Relationship between above-ground biomass and basal stem area.\\
\textbf{Fig. S5} Plant functional types diverge strongly in mass-based partitioning, but converge in area-based partitioning.\\
\textbf{Notes S1} Modelling of biomass partitioning in global vegetation models (GVMs)

### 2015-07-21 manuscript_suppinfo.Rnw
More writing!; Formatting is now complete, I think.

\subsection{Sample sizes by plant functional type}

@@

\clearpage
\subsection{Supporting figures}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureS1.pdf}
    \caption{Global coverage of the climate space by the dataset. Methods are as in Fig. 1 in main text, symbols are colored by vegetation type as reported by the data contributors.}
\label{fig:figureS1}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureS4.pdf}
    \caption{Relationship between above-ground biomass and basal stem area, each point is an individual plant. Lines are spline fits. The right panel shows least-square means of the ratio of above-ground biomass to basal stem area, estimated from a linear mixed-effects model with basal stem area (linear and squared terms) and PFT. }
    \label{fig:figureS4}
\end{figure}

@@

\clearpage
\subsection{Modelling of biomass partitioning in global vegetation models (GVMs)}
\label{sec:gvmsummary}

>>>>

@@

\clearpage

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureS1.pdf}
    \caption{Global coverage of the climate space by the dataset, labelled by vegetation type. Methods for climate variables are as in Fig. 1 in main text. Symbols are colored by vegetation type as reported by the data contributors.}
\label{fig:figureS1}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureS4.pdf}
    \caption{Relationship between above-ground biomass and basal stem area. Each point is an individual plant. Lines are spline fits. The right panel shows least-square means of the ratio of above-ground biomass to basal stem area, estimated from a linear mixed-effects model with basal stem area (linear and squared terms) and PFT. }
    \label{fig:figureS4}
\end{figure}

@@

\clearpage
\subsection*{Notes S1. Modelling of biomass partitioning in global vegetation models (GVMs)}
